Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-12-07 paper.tex
Update constrained extend algorithm.

\textbf{We show here how a multiple constraint solver works: its
  purpose is to find the root $q$ of a non-linear $C^1$ function
  $f(q)$ with a tolerance of $\varepsilon$. If we want to find a
  configuration on a manifold $\manifold$, $f(q)$ can be defined as a
  vector valued function that contains the concatenation of all
  constraints defining $\manifold$. Note that as the intersection of
  two or more manifolds is also a manifold, this constraint solver
  allows us also to generate configurations that lie at the
  intersection of several manifolds.}

@@

\textbf{We observed that values of $\varepsilon=10^{-6}$,
  $\alpha=0.1$, $\alpha_{max}=0.95$ and $w=0.8$ lead to good behavior,
  i.e. fast convergence and low failure rate. These values are kept
  constant for all scenarios in this work.}

\begin{algorithm}
\caption{\texttt{SolveConstraints}($q$, f, $\varepsilon$): find $q$
  such that $f(q) = 0$}
\label{algo:newton}
\begin{algorithmic}
\STATE $i=0$
\WHILE{$\|f(q)\| > \varepsilon$ and $i\leq it_{max}$}
\STATE {\color{red} // $\left(.\right)^{+}$ denotes the Moore-Penrose pseudo-inverse}
\STATE $q \leftarrow$ $q - \alpha \left(\frac{\partial f}{\partial q}(q)\right)^{+} f(q)$
\STATE $i$ $\leftarrow$ $i+1$
\STATE {\color{red}// Make $\alpha$ tend toward $\alpha_{max}$}
\STATE $\alpha \rightarrow \alpha_{max} - w(\alpha_{max} - \alpha)$
\ENDWHILE
\IF {$\|f(q)\| \leq \varepsilon$}
\STATE return $q$
\ELSE
\STATE return failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[h]
\centering
\begin{minipage}[c]{0.6\linewidth}
\begin{tikzpicture}[y=0.55pt, x=0.55pt,yscale=-1, inner sep=0pt, outer sep=0pt]
\definecolor{dg}{rgb}{0,0.3,0}
\path[draw=black,line join=miter,line cap=butt,line width=0.800pt, fill=gray]
  (231.3249,143.2249) .. controls (178.0814,153.1889) and (178.5527,172.4247) ..
  (180.8173,200.8036) .. controls (183.2959,231.8645) and (195.6610,188.6038) ..
  (256.5787,230.0980) .. controls (295.8993,256.8813) and (346.4296,308.3911) ..
  (295.9747,178.5802) .. controls (275.5517,126.0356) and (304.8066,129.4735) ..
  (231.3249,143.2249) -- cycle;
\node at (230,180) {Obstacle};
\path[dashed,draw=dg,line join=miter,line cap=butt,line width=1.5pt]
  (88.8934,277.5752) .. controls (196.3490,276.5363) and (532.8532,213.7494) ..
(434.3656,81.6056) node [above = 0.1cm, color=dg, text width=1.8cm] {\small{Constrained manifold $\manifold$}};
\node [draw,circle,inner sep=2pt,fill = gray] (qn) at (97,277) {};
\node [above = 0.15cm] at (qn) {$q_{near}$};
\node [draw,circle,inner sep=2pt] (qr1) at (367,99) {};
\node [left = 0.2cm] at (qr1) {$q_{rand}$};
\draw [dashed,thin] (qn) -- (450,277);
\node [draw,circle,inner sep=2pt] (qr2) at (367,277) {};
\node [below right = 0.1cm] at (qr2) {$q_{rand}'$};
\node [draw,circle,inner sep=2pt,fill = gray] (qr3) at (345,225) {};
\node [above = 0.22cm] at (qr3) {$q_{rand}''$} ;
\draw [dashed,thin] (qr1) -- (qr2) -- (qr3);
\node [draw,circle,inner sep=2pt,fill = gray] (qnew) at (270,250) {};
\node [above = 0.1cm, left = 0.2cm] at (qnew) {$q_{new}$};
%\draw [thick] (qn) -- (qnew);

@@

\caption{\textbf{One step of constrained extension illustrating
    Algorithm \ref{alg:constrained}: $q_{rand}$ is first projected on
    the tangent space of $\manifold$. $q_{rand}'$ is then projected
    onto $\manifold$, and a classic RRT extension tries to go as far
    as possible from $q_{near}$ towards $q_{rand}''$ while remaining
    on $\manifold$. $q_{new}$ is then returned.}}
\label{fig:gikrrt}
\end{figure}

\begin{algorithm}[h]
  \caption{\texttt{ConstrainedExtend}($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE {\color{red} // $I$ denotes the identity matrix.}
    \STATE $q_{rand}' \leftarrow q_{near} +   
    \left(I -  
    \left(\frac{\partial f}{\partial q}(q_{near})\right)^{+}
    \frac{\partial f}{\partial q}(q_{near})\right) (q_{rand} - q_{near})$
    \IF {$\texttt{SolveConstraints}(q_{rand}',f,\epsilon) = q_{rand}''$}
    \STATE RRT::Extend($\mathcal{T},q_{near},q_{rand}''$)
    \ENDIF
  \end{algorithmic}
\end{algorithm}

>>>>

\textbf{We show here how a multiple constraint solver works: its
  purpose is to find the root $q$ of a non-linear $C^1$ function
  $f(q)$ with a tolerance of $\epsilon$. If we want to find a
  configuration on a manifold $\manifold$, $f(q)$ can be defined as a
  vector valued function that contains the concatenation of all
  constraints defining $\manifold$. Note that as the intersection of
  two or more manifolds is also a manifold, this constraint solver
  allows us also to generate configurations that lie at the
  intersection of several manifolds.}

@@

\textbf{We observed that values of $\epsilon=10^{-6}$,
  $\alpha=0.1$, $\alpha_{max}=0.95$ and $w=0.8$ lead to good behavior,
  i.e. fast convergence and low failure rate. These values are kept
  constant for all scenarios in this work.}

\begin{algorithm}
\caption{\texttt{SolveConstraints}($q$, f, $\epsilon$): find $q$
  such that $f(q) = 0$}
\label{algo:newton}
\begin{algorithmic}
\STATE $i=0$
\WHILE{$\|f(q)\| > \epsilon$ and $i\leq it_{max}$}
\STATE {\color{red} // $\left(.\right)^{+}$ denotes the Moore-Penrose pseudo-inverse}
\STATE $q \leftarrow$ $q - \alpha \left(\frac{\partial f}{\partial q}(q)\right)^{+} f(q)$
\STATE $i$ $\leftarrow$ $i+1$
\STATE {\color{red}// Make $\alpha$ tend toward $\alpha_{max}$}
\STATE $\alpha \rightarrow \alpha_{max} - w(\alpha_{max} - \alpha)$
\ENDWHILE
\IF {$\|f(q)\| \leq \epsilon$}
\STATE return $q$
\ELSE
\STATE return failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[h]
\centering
\begin{minipage}[c]{0.6\linewidth}
\begin{tikzpicture}[y=0.55pt, x=0.55pt,yscale=-1, inner sep=0pt, outer sep=0pt]
\definecolor{dg}{rgb}{0,0.3,0}
\path[draw=black,line join=miter,line cap=butt,line width=0.800pt, fill=gray]
  (231.3249,143.2249) .. controls (178.0814,153.1889) and (178.5527,172.4247) ..
  (180.8173,200.8036) .. controls (183.2959,231.8645) and (195.6610,188.6038) ..
  (256.5787,230.0980) .. controls (295.8993,256.8813) and (346.4296,308.3911) ..
  (295.9747,178.5802) .. controls (275.5517,126.0356) and (304.8066,129.4735) ..
  (231.3249,143.2249) -- cycle;
\node at (230,180) {Obstacle};
\path[dashed,draw=dg,line join=miter,line cap=butt,line width=1.5pt]
  (88.8934,277.5752) .. controls (196.3490,276.5363) and (532.8532,213.7494) ..
(434.3656,81.6056) node [above = 0.1cm, color=dg, text width=1.8cm] {\small{Constrained manifold $\manifold$}};
\node [draw,circle,inner sep=2pt,fill = gray] (qn) at (97,277) {};
\node [above = 0.15cm] at (qn) {$q_{near}$};
\node [draw,circle,inner sep=2pt] (qr1) at (367,99) {};
\node [left = 0.2cm] at (qr1) {$q_{rand}$};
\draw [dashed,thin] (qn) -- (450,277);
\node [right = 0.2cm] at (450,277) {\small{$T_q\manifold$}};
\node [draw,circle,inner sep=2pt] (qr2) at (367,277) {};
\node [below right = 0.1cm] at (qr2) {$q_{rand}'$};
\node [draw,circle,inner sep=2pt,fill = gray] (qr3) at (345,225) {};
\node [above = 0.22cm] at (qr3) {$q_{rand}''$} ;
\draw [dashed,thin] (qr1) -- (qr2) -- (qr3);
\node [draw,circle,inner sep=2pt,fill = gray] (qnew) at (270,250) {};
\node [above = 0.1cm, left = 0.2cm] at (qnew) {$q_{new}$};
%\draw [thick] (qn) -- (qnew);

@@

\caption{\textbf{One step of constrained extension illustrating
    Algorithm \ref{alg:constrained}: $q_{rand}$ is first projected on
    $T_q\manifold$ the tangent space of $\manifold$. $q_{rand}'$ is
    then projected onto $\manifold$. A classic RRT extension tries to
    go as far as possible from $q_{near}$ towards $q_{rand}''$ while
    remaining on $\manifold$. $q_{new}$ is then returned.}}
\label{fig:gikrrt}
\end{figure}

\begin{algorithm}[h]
  \caption{\texttt{ConstrainedExtend}($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE $d \leftarrow$ \texttt{Distance}($q_{near}, q_{rand}$)
    \STATE $q \leftarrow$ $q_{near}$
    \WHILE{$d > \epsilon$}
    \STATE $q_{rand}' \leftarrow$ \texttt{OrthogonalProject}($q_{rand}, T_q\manifold$)
    \STATE $q_{rand}'' \leftarrow$ \texttt{SolveConstraints}($q_{rand}',f,\epsilon$)
    \STATE $d \leftarrow$ \texttt{Distance}($q,q_{rand}''$)
    \STATE $q \leftarrow q_{rand}''$
    \ENDWHILE
    \STATE $q_{new} \leftarrow$ RRT::Extend($\mathcal{T},q_{near},q_{rand}''$)
  \end{algorithmic}
\end{algorithm}
